---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# brazilmaps

`brazilmaps` provides simplified Brazilian territorial meshes as local `sf`
objects. Spatial files are installed with the package: using a map never
triggers a download and does not require an internet connection.

Municipal meshes are bundled at milestones in the official IBGE archive.
Within each consecutive run with the same number of municipalities, the
package retains only the latest edition:

```{r inventory}
library(brazilmaps)
brmap_editions()[c("year", "n_features", "file_bytes")]
```

The current edition is the default. Select a historical edition with `year`:

```{r historical-example, fig.width=7, fig.height=3.5}
current <- get_brmap("municipality", filters = list(state = 50))
mesh_2000 <- get_brmap(
  "municipality", year = 2000, filters = list(state = 50)
)

plot_brmap(mesh_2000) +
  ggplot2::labs(title = "Municipal mesh of Mato Grosso do Sul — 2000")
```

The meshes are simplified for thematic maps, exploratory analysis and joins.
They must not be used for cadastral work, legal boundary decisions or precise
area/perimeter measurement.

## Installation

Install the released version from CRAN:

```r
install.packages("brazilmaps")
```

Or install the development version:

```r
# install.packages("pak")
pak::pak("rpradosiqueira/brazilmaps")
```

## Geographic levels

```{r levels}
states <- get_brmap("state")
immediate <- get_brmap(
  "immediate_region", filters = list(state = 26)
)
intermediary <- get_brmap("intermediate_region")
```

Available levels are Brazil, regions, states, immediate regions, intermediate
regions and municipalities. The package also retains:

- `microregion` and `mesoregion`, discontinued IBGE statistical divisions,
  for compatibility with historical data;
- `state_hex` and `state_region`, curated equal-shape state cartograms.

Earlier PascalCase level names remain accepted with deprecation warnings.
Returned spatial columns use explicit lower-snake-case names such as
`municipality_code`, `state_code`, `immediate_region_code` and
`state_abbreviation`.

## Filtering, joining and plotting

All filter conditions are combined with logical AND:

```{r join-plot, fig.width=7, fig.height=4}
data("gini2015")

midwest_states <- get_brmap(
  "state",
  filters = list(region = 5)
)

plot_brmap(
  midwest_states,
  data = gini2015,
  by = c("state_code" = "cod"),
  fill_by = "gini"
) +
  ggplot2::labs(title = "Gini index — 2015")
```

`join_brmap()` preserves `sf`, and `plot_brmap()` returns a regular `ggplot`
that can be extended with ggplot2 layers and scales.

## Territorial codes

```{r dtb}
get_dtb(code = c(1, 50, 5101837))
head(get_dtb_levels(
  c("municipality", "state"),
  filters = list(state = 50)
))
```

The installed DTB contains the current hierarchy, including 5,571 municipal
units, 510 immediate regions and 133 intermediate regions. When a historical
municipal map is filtered by an immediate or intermediate region, the current
hierarchy is used for municipality codes that still exist.

## Data provenance and maintenance

Municipal geometries come from the official
[IBGE municipal mesh archive](https://geoftp.ibge.gov.br/organizacao_do_territorio/malhas_territoriais/malhas_municipais/).
Current hierarchy data come from the official
[IBGE Localities API](https://servicodados.ibge.gov.br/api/docs/localidades).
All installed geometry uses SIRGAS 2000 (EPSG:4674).

Bundled data retain attribution to their source institutions. Package code is
distributed under GPL-3; detailed data provenance and attribution are recorded
in `inst/COPYRIGHTS`.

Downloads occur only in the maintainer scripts under `data-raw/`. Those scripts
standardize, simplify, rebuild topology, validate counts and codes, test planar
and spherical validity, and record checksums and exact source URLs. Package
users only read the resulting local files.
