---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# brazilmaps

The goal of brazilmaps is to provide Brazilian map spatial objects of varying region types (e.g. cities, states, microregions, mesoregions).

## Installation

Install the release version from CRAN:

```{r , eval = FALSE}
install.packages("brazilmaps")
```

or the development version from github

```{r , eval = FALSE}
# install.packages("devtools")
devtools::install_github("rpradosiqueira/brazilmaps")
```

## Example

Let's assume that we want to plot the brazilian municipalities of the Midwest Region. To do this simply execute:

```{r midwest-cities, eval=FALSE}
library(brazilmaps)

# Get de map
midwest_cities <- get_brmap(geo = "City",
                            geo.filter = list(Region = 5))

# Plot
plot_brmap(midwest_cities)

```

I'm preparing a vignette with more examples that will come soon.
